Towards a standardized evaluation of imputation methodology

### 2022-09-19 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Restyled some paragraphs

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some overarching issues with assessing imputation method performance. In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation (e.g. focusing on one performance measure over another, may yield different conclusions about imputation efficacy). In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results. Identifying problems with the imputation-generating process (e.g., an iterative imputation algorithm) may offer explanations for under-performance of imputation methods. Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked. %(i.e. the complete-data model used to estimate the comparative truth). % The choice of performance measures might inadvertently influence simulation results.

@@

(i) The estimates should preferably be unbiased. Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero). In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.

(ii) The estimates should have a proper coverage rate. Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient. Under-coverage (when estimation is too liberal and intervals are too narrow) indicates that the procedure is not confidence valid and may lead to invalid inference. Over-coverage (when estimation is too conservative and intervals are too wide) tells us that efficiency could still be gained.

@@

(iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods. Wider intervals are associated with more uncertainty whereas a more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid than a more narrow interval that is not properly covered anymore.

(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[][]{buur18}. However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.

If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified. To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.

>>>>

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some overarching issues with assessing imputation method performance. In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation. For example, focusing on one performance measure over another, may yield different conclusions about imputation efficacy. In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results. Identifying problems with the imputation-generating process (e.g., an iterative imputation algorithm) may offer explanations for under-performance of imputation methods. Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked. %(i.e. the complete-data model used to estimate the comparative truth). % The choice of performance measures might inadvertently influence simulation results.

@@

(i) The estimates should preferably be unbiased. Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results. A negligible absolute bias, for example, may already yield an almost infinite relative bias if the true value of the estimand is zero. In most cases, unbiased estimates may be expected under a MCAR missingness mechanism which can be easily verified by including MCAR missingness to the scope of missingness investigations.

(ii) The estimates should have a proper coverage rate. Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep{neym34}. Under-coverage indicates that the procedure is not confidence valid and may lead to invalid inference. Undercoverage may occur when the estimation procedure is either biased, too liberal leading to intervals that are too narrow, or both. Over-coverage, though technically not confidence invalid, would indicate that efficiency could still be gained. In such cases, a narrower interval exists for which the method would still yield confidence valid result.

@@

(iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods. Wider intervals are associated with more uncertainty whereas a more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid than a more narrow interval that is not properly covered anymore. That said, with valid nominal coverage, the method that yields the narrowest interval would be most efficient.

(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[Chapter 2.6][]{buur18}. However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.

If the goal is inference --not prediction-- then the uncertainty about estimates needs to be properly quantified. To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.

GV CONTINUE HERE
